\chapter[Discussion \& Outlook]{Discussion and Outlook}\label{ch:discussion}

\section{Discussion}
In this thesis we presented the application of the fast position readout of an 
optically trapped spherical particle to two micro scale acoustofluidics (MSAF) 
phenomena: 1) the transient build up behaviour of acoustic streaming (AS) as 
well as of the acoustic radiation force (ARF); 2) the viscous torque induced 
rotation of spherical particles in a viscous fluid where the viscous boundary 
layer (VBL) thickness is as large as the particle radius itself $\delta \approx 
R$.

The investigations of both phenomena were possible because the laser beam 
causing the trapping potential carries positional information about the 
relative particle position with respect to the laser focal point. In order to 
extract this information from the laser beam, the beam has to be collimated 
again after the trapping and focused onto photo detectors. For the resolution 
of two orthogonal directions of the in-plane movement, the photo detector (PD) 
has to be a quadrant photo detector (QPD). For the axial movement a single PD 
is sufficient because only the total intensity onto the PD is needed. However, 
we have two QPDs in our setup.

The physical unit of the QPDs output is \si{\volt}. In order to convert the 
voltage to the unit of \si{\meter}, we took advantage of the linear regime of 
QPDs where a change in measured voltage is proportional to the movement of the 
particle within the trap. In this regime, the optical trap (OT) can also be 
used as force measurement device because it has the same properties as a linear 
mechanical spring for small particle displacements in the order of $0.1\R$. 
The respective start and end point of this regime is dependent on a multitude 
of parameters. But, movements of less than \SI{100}{\nm} are still linear for 
our OT setup. The voltage-conversion factor as well as the stiffness of the OT 
can be derived from the single-sided frequency spectrum of the trapped 
particle. The particles in the fluid suspended were so lightweight and small 
that they showed visible Brownian motion. Brownian motion is a random process 
with the property that over a long time period the particle will be back at its 
initial position. Also, while being trapped the particle showed Brownian 
motion. The frequency content of a trapped particle in any spatial direction 
follows the curve of a low-pass filter. With the amplitude at zero frequency 
$A_{0}$ and the cutoff frequency $f_{\MR{c}}$ where 
$A\vert_{f=f_{\MR{c}}}=\sfrac{A_{0}}{2}$ it was possible to calculate the 
voltage-meter conversion factor as well as the stiffness for each spatial 
direction independently.

Besides the size of the trapped particle the conversion factor is dependent on 
the viscosity of the surrounding fluid. Exact knowledge of the viscosity is a 
requirement for quantitatively precise measurements. The viscosity of the fluid 
is in general a function of its temperature. While the ambient temperature was 
straightforward to measure, the measurement of the temperature within a MSAF 
device or at the focal point of the trapping laser was cumbersome. Although the 
intensity of our laser exceeds the sun's intensity by orders of magnitudes, we 
showed in \cref{sec:TO-temperature} that for our setup the laser induced 
temperature change and therefore the temperature induced viscosity change is 
negligible for all our experiments. However, this is not true for every OT 
because the temperature change does not only depend on the laser power itself 
but even more on the particle material and the laser wavelength-fluid 
combination, and therefore the absorption at a specific wavelength.

To investigate the first MSAF phenomena, we developed a new measurement method 
in \cref{ch:timeconstant} to visualize the transient build up of the ARF and AS 
and then studied in \cref{ch:pulsing} the effects of a pulsed excitation on the 
respective build up times. Key for those experiments was, that the ARF and AS 
induced particle displacement were along orthogonal spatial direction. 
Moreover, the ARF displacement was along one of the axes of the in-plane QPD 
and the AS displacement was along the axial direction on another QPD. Hence, 
the measured displacement data is also separated and cross-talk free. We 
validated the orthogonality of the ARF and the drag force from AS by steady 
state force measurements with two different sized particles where all 
experimental parameters besides the particle radius were kept the same. The 
drag force from AS scales with the particle radius, $\FAS\propto R$ and the ARF 
with the volume of the particle, $\FARF\propto R^{3}$. The ratio of the 
measured forces along same directions but of different particle sizes showed 
the same scaling laws. 

However, without modifications on the OT we were not able to measure any of 
those effects because the timeconstant of the OT $\tOT$ is much larger than the 
theoretical build up time of the ARF. In fact, for our set of experimental 
parameters and our device configuration $\tOT \approx \tas \gg \tarf$ with 
$\tas = \SI{1.59}{\ms}$ and $\tarf = \SI{1.4}{\us}$, respectively. We reduced 
$\tOT$ effectively to zero by reducing the laser power to almost zero such that 
there was no effective trapping by the OT. The laser cannot be switched off 
completely because the beam is the cause for the measured intensity at any of 
the QPDs.

Because of the laser power reduction, we also needed to install an optical 
shutter right before the QPDs. The laser intensity is too high for the photo 
detectors. In normal trapping mode a set of filters reduces the intensity to 
less than 0.01\% of the incoming intensity. The optical filter is actuated to 
be almost completely closed such that it acts as intensity filter if the laser 
power is high (trapping) and opens as soon as the laser power is reduced for 
the measurement itself.

The laser power change occurred almost instantaneously but the shutter had an 
opening time of less than \SI{15}{\ms}. To ensure undisturbed data we waited 
\SI{25}{\ms} before starting the acoustic excitation. During this time the 
particle will start to sediment due to gravity because it was not trapped 
stably anymore. However, this movement was only $\approx 0.05\,R$ along the 
direction of the laser beam and, hence, did not hinder the measurement; the 
in-plane movement was zero because no forces besides gravity are acting on the 
particle.

The (pulsed) acoustic excitation was then switched on for only \SI{30}{\ms} 
because a longer excitation would lead to such large displacements -- primarily 
along the ARF direction -- that the same particle could not be re-trapped by 
the OT. For our studies we were not interested in the actual displacements but 
in the time it takes until the movement along any direction starts. For actual 
displacement magnitudes one would need to retrieve the voltage-meter conversion 
factor but for that the particle needs to be trapped. Hence, only the voltage 
at the QPDs can be investigated.

Because the particle is free-floating and, therefore, uncontrolled in its 
location for the first \SI{25}{\ms} we repeated the measurement multiple times 
per location and averaged the measured data. To further subtract the gravity 
induced movement from the data we perform the same amount of measurements with 
no acoustics at all and then subtract the averaged \emph{no-acoustic} data from 
the \emph{acoustics} data.

The data for both types of experiments (non-pulsed and pulsed) showed that the 
ARF induced displacement starts immediately whereas the AS induced displacement 
takes significantly longer. The measured time offset between ARF and AS is an 
explanation why in experiments (\cname{Hoyos2013} and~\cname{Castro2016}) a 
pulsed acoustic excitation suppressed the build up of AS whereas numerical 
investigations~\cname{Muller2015} with an ideal fluid cavity revealed a much 
smaller offset which did not suppress streaming. Although the data had the unit 
of \si{\volt} and not \si{\meter}, the slope of the AS and ARF induced 
displacement was showing that the acceleration after the build up by any of two 
effects occurs fast and then transitions into constant velocity (linear slope).  
A constant velocity implies the equilibrium between the force pulling on the 
particle and the hindering drag force.

% It is a valid statement because the measured voltage (within the linear regime 
% of the QPDs) is the displacement divided by the voltage-meter conversion 
% factor. The slope of the voltage-over-time data equals the slope of the 
% displacement which is in turn the velocity of the particle.

Additionally, the pulsed excitation measurements revealed that a pulsed 
excitation leads to a greater reduction in terms of the final displacement at 
the end of the measurement on the AS than it did on the ARF. We measured this 
observation for all pulsing frequencies. The ARF and AS are linearly dependent 
on the acoustic energy density within the fluid cavity. All our data also 
showed this relation but with a steeper slope for AS than the ARF. The pulsed 
excitation results help to further explain the experimentally shown suppression 
of AS~\cite{Hoyos2013,Castro2016}. However, with our measurement protocol we 
can only measure for the first \SI{30}{\ms} of the acoustic excitation, but any 
MSAF application is operated continuously where all fields have developed 
fully.
% At the moment, we cannot measure these effects during steady-state.


Compared to the first phenomena, no setup modifications at the OT were 
necessary for the second MSAF phenomenon: the rotational speed measurement of a 
spherical particle driven by the viscous torque (\cref{ch:viscoustorque}). The 
two main requirements for the generation of a viscous torque that is 
sufficiently large to initiate a particle rotation are: 1) the formation of a 
VBL $\delta$ around the particle; 2) a two dimensional orthogonal acoustic 
excitation with phase difference such that the acoustic pressure at the 
particle surface has a continuous phase change in circumferential direction.

The second requirement was straightforward to accomplish by the use of a 
quadratic fluid cavity where the excitation is applied to the system from two 
orthogonal directions. The phase difference of the two excitation signals could 
easily be realized with most dual channel function generators. The viscosity 
increase of the fluid and therefore the larger VBL around the particle can be 
created by a glycerol water mixture (3:7 for our experiments). A greater 
viscosity increase was not possible, because glycerol has the side effect that 
it also increases the refractive index of the mixture. For stable trapping the 
refractive index of the particle must be larger than of the surrounding medium. 
For equal refractive indices the particle is \emph{invisible} for the laser 
beam because no path change of the light occurs at the particle-fluid 
interface. A greater index of refraction of the particle than the fluid index 
leads to repulsion of the particle from the laser focal point.

Our measurement utilized the property that a distinct amplitude peak appears at 
the rotational frequency of the particle in the one sided power spectrum of the 
trapped rotating particle. We validated this finding with deformed and partial 
gold coated particles where the rotation was also visible optically. By having 
a highspeed camera mounted to the OT and choosing the excitation parameters 
such that the viscous torque induced rotation is clearly visible, the 
rotational speed measurement from the power spectrum could be validated by the 
analysis of the single frames from the video. The non-spherical particles did 
not rotate with a single constant velocity because of their non-spherical 
shape.  Therefore, multiple additional peaks were also visible in the power 
spectrum.  However, the partially gold coated particles with perfect spherical 
shape showed a single peak that matched the rotational speed from the optical 
analysis.

With a validated method for measuring the rotational speed of a spherical 
particle we investigated the situation where the particle radius is as big as 
the VBL thickness itself ($R\approx\delta$). In the regime where $\delta > 
\sfrac{R}{15}$ a numerical study by \cname{Hahn2016} predicted the final 
terminal rotational velocity order of magnitudes lower than a theoretical 
investigation by \cname{Lamprecht2015}. Note here, that~\cname{Lamprecht2015} 
restricted their calculations to the assumption that $R\gg\delta$. Our 
measurements of rotations up to \SI{13700}{\rpm} confirmed the results from 
\cname{Hahn2016} and also confirmed the limited validity of the theoretical 
formula of \cname{Lamprecht2015} in the regime where $R\approx\delta$.

\section{Outlook}

The OT is a great tool for measuring acoustic effects on micro-scale particles 
suspended in a fluid. The four main advantages are: 1) measurements on a single 
particles; 2) high spatial precision and repeatability of the measurement 
locations; 3) high temporal resolution of particle position; d) high 
sensitivity to surroundings. The last point is also by far the biggest 
disadvantage. For all our experiments we were very cautious about us being in 
the lab during the measurements and any other disturbances, like fast and 
careless opening/closing of the door or like constructions works in the 
neighboring lab. In any case, the combination of the optical trap as 
characterization/measurement device for the acoustical trap offers many 
possible future research contributions. In the following we want to categorize 
our suggestions in three sections: a) deeper theoretical understanding of the 
OT; b) future experiments with our setup where no big modifications are needed; 
c) more general ideas which would need larger modifications to our setup.

\subsection{Optical Trap Calibration}

In this thesis we did not explain the calibration protocol for our OT. As 
before, interested readers are pointed to the thesis of \cname{Lamprecht2017} 
(Section 3.2.4) where our calibration process is explained in great detail. The 
basic procedure is to fit a low-pass filter to the frequency content of a 
trapped particle. This fitting is based on two parameters: the amplitude 
$A_{0}$ at zero frequency and the cut off frequency $\cutoff$. With those two 
parameters the stiffness $\kappa_{i}\propto\cutoff$ and the voltage-meter 
conversion factor $\Upsilon_{i}\propto \sfrac{1}{(\cutoff\sqrt{A_{0}} )}$ can 
be calculated for all three spatial directions separately. The total force is 
then simply the product of the measured voltage times the two factors and, 
hence, \begin{equation}
  \force_{i,\MR{measured}}\propto \kappa_{i}\,\Upsilon_{i}\propto 
  \sfrac{1}{\sqrt{A_{0}}}.
\end{equation}
Interestingly, the measured force is only proportional to one of those fitting 
parameters $A_{0}$. The fit itself, especially, the parameter $A_{0}$ depends 
highly on the beginning of the data range used for the fit. From experience we 
used for our OT calibration the frequency spectrum data ranging from 
\SI{20}{\hertz} up to \SI{6}{\kilo\hertz} for the $\ex$ and $\ey$ direction and 
\SI{10}{\hertz} to \SI{2}{\kilo\hertz} for $\ez$. The $\ez$ direction has a 
smaller and lower range because it is known that OTs are in general weaker 
along the beam axis and, hence, the respective cut-off frequency is also lower. 
Therefore, an investigation regarding the fitting interval will help increase 
the absolute precision of OTs.

However, the error introduced by a false data range is \emph{just} a false 
scaling of the forces. Measurements that use the same fitting procedure can 
still be compared because all measurements might share the same scaling error. 
Up to our knowledge we are not aware of any publication that discusses (and 
resolves) this small but important detail for the power spectrum OT 
calibration.

\subsection{Future Research with our Optical Trapping Setup}

Although we investigated two MSAF phenomena with our OT setup, there are many 
more open questions and interesting projects. We select three projects which we 
think are of great interest for the MSAF community and also achievable without 
any modifications to our present OT setup.

The first is a variation of \cref{ch:timeconstant,ch:pulsing}. In a first step 
we were able to measure the build up of the two effects, but only for the first 
120'000 excitation periods ($\approx \SI{30}{\ms}$). Although we could show 
that AS is significantly slower in its build up than the ARF, we could not 
investigate the steady-state behavior of a pulsed excitation. Therefore, we 
suggest to measure the steady-state forces exerted on a single particle for 
various pulse frequencies settings and compare the forces to a non-pulsed 
measurement. To reduce uncertainties between different settings and increase 
data comparability, one should create a measurement protocol where all data is 
measured with one device initialization.

The second possible field of interest is also investigating the ARF and AS. 
However this time, one would be investigating the scaling laws of AS and the 
ARF. From theory we know that they scale $\propto R$ and $\propto R^{3}$ for a 
spherical particle, respectively. The total acoustic force 
$\forcevector^{\MR{ac}}$ is understood to contain a linear part from AS and a 
cubic part due to the ARF
\begin{equation}
  \forcevector^{\MR{ac}}\left( R \right) = \vb{C}^{\MR{AS}}\,R + 
  \vb{C}^{\MR{rad}}\,R^{3}.
\end{equation}
From force measurements of two different sized particles with the same 
experimental condition it is possible to extract the constants 
$\vb{C}^{\MR{AS}}$ and $\vb{C}^{\MR{rad}}$. Therefore one could image the pure 
fields that are creating the ARF and also that are due to AS. In order to 
validate the assumed composition of the total acoustic force on a particle 
holds, a third measurement with another particle size is needed. Based on the 
constants $\vb{C}^{\MR{AS}}$ and $\vb{C}^{\MR{rad}}$ one is able to predict the 
force field in all three dimensions for the third particle size $R_{3}$
\begin{equation}
  \forcevector^{\MR{ac}}\left( R_{3} \right) = \vb{C}^{\MR{AS}}\,R_{3} + 
  \vb{C}^{\MR{rad}}\,R^{3}_{3}.
\end{equation}
The main challenge for this study is to ensure the same pressure amplitude 
inside the cavity for the three different particle sizes.

The third and last potential study we want to discuss here is a controversy 
about the ARF for heavy particles in viscous fluids. \cname{Baasch2019} showed 
numerically that the contributions of microstreaming around heavy particles in 
a viscous fluid are of such significance that they could lead to a 
force-inversion. Their findings strengthen the theory of 
\cname{Doinikov1994Rigid} which also predicts this force inversion whereas the 
theory of \cname{Settnes2012} does not. With the measurement of the rotational 
speed in a glycerol-water mixture we could already show that the OT can operate 
within viscous fluid. Besides being heavy the other main requirement is to be 
trappable for the OT setup. Additionally, those particles should have the same 
or smaller optical absorption as the surrounding fluid because otherwise the 
laser induced heating might be too high such that the changes in viscosity are 
not negligible anymore.

\subsection{Future Optical Trap Research for MSAF}

In the final part of this thesis we will motivate three possible studies that 
are not applicable to our specific OT setup. For the first one, the main 
limitation is not directly our OT but the devices we use. The main requirement 
for our devices is that they are transparent from top to bottom for our laser 
light as well as being wide enough to accommodate the converging cone of the 
laser beam. Therefore, we cannot measure close to walls because the silicon 
walls are non-transparent for our laser wavelength and hinder the laser beam 
and, hence, would violate the assumption of a symmetrical trapping potential 
which is the basis for the trap calibration. Acoustic wall effects are a 
interesting phenomena that are important for narrow channels and could lead to 
further applications. Devices where also the channel side walls are made out of 
glass (all-glass devices) might be an interesting opportunity for studying 
these effects. However, one would need to investigate the refractive index of 
the fluid and the specific glass carefully.

For our investigated applications this is not a problem because the two 
interfaces of the top cover glass with the immersion oil layer and the fluid 
inside the cavity are parallel. Therefore, the beam passes this interface 
without a change in direction. However, in an all-glass device which would be 
used for measuring close to the channel wall some parts of the laser beam will 
be going through the channel wall. For different refractive indices of the 
fluid and the glass this will lead to a deflection of the laser which in turn 
will violate the assumption of a symmetric optical potential. The magnitude of 
the violation is increasing with the index difference. For matching indices of 
the device material and the fluid no deflection at any of those interfaces 
occurs.

Secondly, with an OT consisting of two lasers where one is creating the optical 
potential for trapping and the other is the light source for the position 
detection at the QPDs one is able to re-perform the build up experiments. 
Besides a validation of our results with another setup and device, one would 
also eliminate the lead time before the acoustic excitation starts. This lead 
time was necessary in our case because we have a single beam OT and needed to 
open the optical shutter before we could start the acoustic excitation.

Lastly, our device is a bulk acoustic wave (BAW) device where a standing 
pressure wave is formed in the bulk of the fluid. The other major kind of 
devices are surface acoustic wave (SAW) devices. In those devices opposing 
traveling waves at the coverplate-fluid interface create a standing pressure 
field inside the fluid that can also manipulate objects. Up to our knowledge no 
direct force measurement were performed with SAW devices so far.
